---
title: "MAS | Aula Prática 4 - Classificador Naïve Bayes "
author: "André Silvestre Nº104532 CDB1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Exemplo sobre dataset `Iris` (com preditores quantitativos)

- Margarida G. M. S. Cardoso
- 17 de fevereiro, 2023

# PL4 Goal:

\textcolor{lightgray}{To use the Naive Bayes classifier considering an example with quantitative predictors.}

- Usar o classificador Naive Bayes considerando um exemplo com preditores quantitativos.

```{r message=FALSE, warning=FALSE}
# Bibliotecas
rm(list=ls(all=TRUE)) # First remove everything! 
library(psych)        # to provide descriptive statistics
library(ggplot2)      # to provide graphics 
library(lsr)          # for Eta
library(e1071)        # classification with Naïve Bayes
library(Metrics)      # used to provide classification metrics
```


\newpage

# 1) The iris data set

Data set by R. Fisher (1936) The use of multiple measurements in taxonomic problems.Annals of Eugenics. 7 (2): 179-188.

### TARGET:

- **Species:** (setosa, versicolor , virginica)

### PREDICTORS: 

- **Sepal.Length** 

- **Sepal.Width**

- **Petal.Length**

- **Petal.Width** 

---

## 1.a)	Get the iris data set

```{r}
data(iris)
names(iris)
dim(iris)
```

## 1.b)	Iris descriptive statistics

```{r}
summary(iris)

#
attach(iris)

#continuing descriptives (with "psych")
describe(iris[,1:4])

# describe predictors for each Species
describe(iris[Species=="setosa",-5])
describe(iris[Species=="versicolor",-5])
describe(iris[Species=="virginica",-5])
```

## 1.c)	Exploring correlations between target and predictors

```{r out.width='48%', fig.show='hold', fig.align='center'}
# Eta_SL
anova_ <- aov(Sepal.Length~ Species, iris) 
Eta_SL <- sqrt(etaSquared(anova_ )[,1])
Eta_SL

# Eta_SW
anova_ <- aov(Sepal.Width ~ Species, iris) 
Eta_SW <- sqrt(etaSquared(anova_ )[,1])
Eta_SW

# Eta_PL
anova_ <- aov(Petal.Length~ Species, iris)
Eta_PL <- sqrt(etaSquared(anova_ )[,1])
Eta_PL

# Eta_PW
anova_ <- aov(Petal.Width ~ Species, iris) 
Eta_PW <- sqrt(etaSquared(anova_ )[,1])
Eta_PW

# correlations between predictors
round(cor(iris[,1:4]),2)

# some graphics
c <- ggplot(iris, aes(x=Petal.Length, fill=Species, color=Species)) + 
geom_histogram(binwidth = 1) + 
labs(title="Petal.Length distribution by species")
c + theme_bw()

#
p <- ggplot(iris, aes(x=Sepal.Width, fill=Species, color=Species)) + 
geom_histogram(binwidth = 0.5) + 
  labs(title="Sepal.Width distribution by species")
p + theme_bw()
```


\newpage

# 2) Performing classification on iris species

## 2.a)	Naive bayes classifier

This classifier assumes independence of predictors (!) and considers $P(C|X_1..X_p)$ is proportional to Produt( $P(X_j|C)) \times P(C)$ (sample estimates are used)

```{r}
x <- iris[,c(2,3)] 
y <- iris$Species
nb.iris<-naiveBayes(x, y) 
nb.iris

#str(nb.iris)
nb.iris$tables$Sepal.Width[1,2]

# Naive Bayes results include, for each numeric predictor, a table giving 
# the mean and standard deviation of the predictor within each target class
```

## 2.b) Naive Bayes based predictions

```{r}
#predicted probabilities
prob.nb_y <- predict(nb.iris,iris,type = "raw") 
head(prob.nb_y)

# predicted classes
pred_y <- predict(nb.iris,iris )  # default is type="class"
head(pred_y)

#
iris_pred<-cbind(iris[,c(2,3,5)],prob.nb_y,pred_y) 
iris_pred[ c(1:3,50:53,100:103),]
```

## 2.c)	Naive bayes performance

```{r}
#
(confusion_mat<-table(iris$Species,pred_y))

#accuracy
acc <- accuracy(iris$Species,pred_y) 
acc

#classification error
ce(iris$Species,pred_y) 

#observations with classification errors
iris_pred[which(iris$Species!=pred_y),]

# majority class frequency
default_p<-max(mean(Species == "setosa"), 
               mean(Species == "virginica"), 
               mean(Species == "versicolor"))
default_p

# Huberty
(Huberty<-(acc-default_p)/(1-default_p))

```


\newpage

# Exercícios

\textcolor{lightgray}{**EXERCISE 1:** considering the nb.iris model, present the calculations of probabilities of belonging to each target class for observation 53 of the iris data set}

- Considerando o modelo nb.iris, apresentar os cálculos das probabilidades de pertencer a cada classe-alvo para observação 53 do conjunto de dados da íris

- Primeiro, precisamos extrair as médias e os desvios-padrão de cada variável preditora para cada classe-alvo. Em seguida, usamos a fórmula do classificador Naive Bayes para calcular as probabilidades de pertencer a cada classe-alvo, dada a observação 53.

```{r}
# selecionar as tabelas de probabilidade correspondentes às variáveis usadas
tabela_sw <- nb.iris$tables$Sepal.Width
tabela_pl <- nb.iris$tables$Petal.Length

# selecionar as observações a serem classificadas
obs_53 <- iris[53, c("Sepal.Width", "Petal.Length")]

# calcular as probabilidades condicionais para cada classe usando a distribuição normal
prob_setosa <- dnorm(obs_53[1], tabela_sw[1, 1], tabela_sw[1, 2]) * dnorm(obs_53[2], tabela_pl[1, 1], tabela_pl[1, 2]) * nb.iris$apriori[1]

prob_versicolor <- dnorm(obs_53[1], tabela_sw[2, 1], tabela_sw[2, 2]) * 
                   dnorm(obs_53[2], tabela_pl[2, 1], tabela_pl[2, 2]) * nb.iris$apriori[2]

prob_virginica <- dnorm(obs_53[1], tabela_sw[3, 1], tabela_sw[3, 2]) * 
                  dnorm(obs_53[2], tabela_pl[3, 1], tabela_pl[3, 2]) * nb.iris$apriori[3]

# normalizar as probabilidades
probs <- c(prob_setosa, prob_versicolor, prob_virginica) / sum(c(prob_setosa, prob_versicolor, prob_virginica))

# exibir as probabilidades de pertencer a cada classe
probs
```


---








\textcolor{lightgray}{**EXERCISE 2:** to evaluate the Naïve Bayes performance on the iris data set, perform train-test approach using a target based stratified sample and 0.65 of observations for training}

- Para avaliar o desempenho da Naïve Bayes no conjunto de dados da íris, realizar uma abordagem de teste de comboio utilizando uma amostra estratificada baseada no alvo e 0,65 de observações para a formação


```{r}
# Load the caret package for data partitioning and model training
library(caret)

# Create a stratified sample to ensure equal representation of each class 
# in both the training and test sets
set.seed(777)
ind_train <- c(sample(1:50,35), sample(51:100,35), sample(101:150,35))
iris_train <- iris[ind_train,] 
iris_test <- iris[-ind_train,] 

# Train the Naïve Bayes model using the training set
nb_model <- train(Species ~ ., data = iris_train, method = "nb")

# Predict the target variable for the test set using the trained model
nb_pred <- predict(nb_model, newdata = iris_test)

# Evaluate the performance of the model using confusion matrix and classification report
confusionMatrix(nb_pred, iris_test$Species)
```

---










\textcolor{lightgray}{**EXERCISE 3:** 
- write the file "iris.csv"; 
- edit this file and remove the Sepal.Length value from observation 1 and the Species value from observation 2; 
- rename this file as $iris_2.csv$ and try to implement the same analysis (i.e. steps $1)$ and $2)$ conducted for iris);
- report the problems found.
Eventually, repeat the analysis after eliminating the rows with missing observations}

```{r}
# - write the file "iris.csv"; 
write.csv(iris, file = "iris.csv", row.names = FALSE)
```


```{r}
# - edit this file and remove the Sepal.Length value from observation 1 
# and the Species value from observation 2; 
# - rename this file as $iris_2.csv$ and try to implement the same analysis 
# (i.e. steps $1)$ and $2)$ conducted for iris);
# - report the problems found.

iris_2 <- read.csv("iris.csv") 
iris_2$Sepal.Length[1] <- NA             # set Sepal.Length in row 1 to NA
iris_2$Species[2] <- NA                  # set Species in row 2 to NA
write.csv(iris_2, file = "iris_2.csv", row.names = FALSE)

# Error in createDataPartition(iris_2$Species, p = 0.65, list = FALSE) : 
#  The data contains more levels than the partitioning variable. 

# Este erro ocorre pq há NAs e ent o R não consegue dividir o dataset
```

```{r}
iris_2 <- read.csv("iris_2.csv")    # read in the modified CSV file
iris_2 <- na.omit(iris_2)           # remove rows with missing values

set.seed(123)
library(caret)
trainIndex <- createDataPartition(iris_2$Species, p = 0.65, list = FALSE)
train <- iris_2[trainIndex, ]
test <- iris_2[-trainIndex, ]
```

